Luminance constancy under fixed geometry

### 2018-05-12 Draft4/draft4.tex
answed some comments

Early work on computational color constancy considered a simplified case with multiple flat matte objects and a single spatially diffuse illuminant \cite{LandRetinex,Buchsbaum80,MaloneyWandell86}.
Subsequent computational work considered more complex geometries and incorporated probabilistic descriptions of the statistics of naturally occurring scenes \cite{funt1988color, D'ZmuraConstancy3, barron2012color, D'ZmuraIversonSinger,BrainardFreeman}.
%LET'S SEE IF WE CAN FIND SOME RECENT REVIEWS).
A key insight from this computational work is that stable color descriptors of a target object cannot be obtained from the light reflected from the target object alone.
But, it is possible to obtain such descriptors by analyzing jointly the light reflected from all of the objects in the scene.
As a consequence, the performance of color constancy algorithms will be affected by variation in the surface reflectance of all of the objects in the scene \cite{BrainardWandellRetinex}, not just by variation in the illuminant; evaluation of constancy algorithms must include assessment of their performance with respect to both variation in the illumination and the surface reflectance of all of the scene objects \cite{BrainardWandellRetinex,BrainardFreeman}.

@@

\subsection{Labeled training data} \label{method:VirtualWorld}
\subsubsection{Virtual naturalistic scenes}
The light that reflects from objects to the eyes depends on many factors.
These include the surface reflectance, texture, material and geometry of the objects, 
the spectral power distribution function of the illuminants, and the position of the observer.
We have developed a rendering package 
(\href{https://github.com/BrainardLab/VirtualWorldColorConstancy}{https://github.com/BrainardLab/VirtualWorldColorConstancy}) 
that allows us to construct models of naturalistic scenes, with key scene factors under programmatic control.
The package builds on our RenderToolblox4 package (\href{http://rendertoolbox.org}{http://rendertoolbox.org}) \cite{heasly2014rendertoolbox3},
and harnesses the open-source computer graphics renderer Mitsuba (\href{https://www.mitsuba-renderer.org}{https://www.mitsuba-renderer.org})
\cite{jakob2015mitsuba} to produce physically-accurate images from the scene models.
Because each image is rendered from a known scene model, each image pixel can be labeled with 
the surface reflectance of the corresponding scene object.
By incorporating statistical models of the variation of object surface reflectance and daylight illumination, 
the pipeline allows us to produce large labeled sets of images that capture key aspects of the task-relevant 
statistical structure of natural scenes.
%% DHB: Need to fix up citations in this paragraph. The URLs for RTB and Mitsuba should be inside the same paraneheses as the citations that
% follow.

@@

For the primary results, we used the ``{\it Library}'' base scene and a spherical target object.
The library base scene contains 2 area lights. 
We inserted one additional spherical spherical light source into the scene.
The position and size of the inserted object, inserted light source and viewpoint were held fixed across all 
scenes in the database. The geometry is shown in Figure~\ref{fig:3DScene}.
Surface and illuminant spectra were drawn according to statistical models of naturally occurring spectra, as described below.
Multispectral images ($320 \times 240$ pixels) were rendered at 31 evenly-spaced wavelengths between $400$nm and $700$nm.
\subsubsection{Illumination spectra}
% Figure 6
\begin{figure}
\centering
    \begin{subfigure}[b]{0.24 \textwidth}
    \centering
	\caption{Granada dataset}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_a.pdf}
        \label{fig:granadaData}
    \end{subfigure}
	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{Statistical model}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_b.pdf}
        \label{fig:illuminantSamples}
    \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{CIE xy chromaticity}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_c.pdf}
        \label{fig:xyDiagram}
        \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{Color swatches}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_d.pdf}
        \label{fig:sRGBIlluminant}
    \end{subfigure}
    \caption{{\bf Statistical model of illumination spectra:} (a) Normalized Granada dataset. Each spectrum is normalized by its mean (with mean taken over wavelength). (b) Sample spectra generated using the statistical model derived from normalized Granada dataset. (c) CIE xy chromaticities of the Granda dataset (black) and the samples shown in b (red). F(d) sRGB renditions of the samples shown in b.}
\label{fig:illuminant}
\end{figure}
%% DHB: The chromaticities in c and renderings in d are for the spectra in b, right?  I wrote the caption to say so, just checking.

@@

Because our multivariate Gaussian model is based on normalized spectra, it does not embody the large variation in overall intensity of natural daylights.
To obtain illuminant spectra with intensity variation similar to the Granada dataset, we scaled each randomly generated relative spectrum.
The scale factors were sampled uniformly from range 0.15 and 150.
%% DHB: Check that these values are right.  Linear spacing or log spacing?

\subsubsection{Surface reflectance spectra}
% Figure 7 Surface Reflectance Data and Model
\begin{figure}
\centering
	\begin{subfigure}{0.24 \textwidth}
    \centering
            \caption{Munsell and Vrhel dataset}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_a.pdf}
        \label{fig:reflectanceSpectra}
    \end{subfigure}
    \begin{subfigure}{0.24 \textwidth}
    \centering
        \caption{Statistical model}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_b.pdf}
        \label{fig:reflectanceSamples}
    \end{subfigure}
    \begin{subfigure}{0.24 \textwidth}
    \centering
    \caption{CIE xy chromaticity}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_c.pdf}
        \label{fig:xyChroReflectance}
    \end{subfigure}    
    \centering
	\begin{subfigure}{0.24 \textwidth}
    \centering
        \caption{Color swatches}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_d.pdf}
        \label{fig:backgroundSwatches}
    \end{subfigure}
    \caption{{\bf Statistical model of surface reflectance:} (a) Surface reflectance spectra from the Munsell and Vrhel datasets. (b) Sample spectra generated using the surface reflectance statistical model. (c) CIE xy chromaticity diagram of the Munsell and Vrhel reflectances and samples shown in b (red). (d) sRGB renditions of the samples shown in b, rendered under the CIE D65 daylight.}
\label{fig:surfaceReflectanceGeneration}
\end{figure}
%% DHB: The chromaticities in c and renderings in d are for the spectra in b, right?  I wrote the caption to say so, just checking.

@@

For generating the target object reflectance at a particular luminance, the values in a generated spectrum were 
scaled such that the LRV had the desired value (see appendix).
Figure \ref{fig:targetSwatches} shows color renderings of target reflectance spectra under CIE illuminant D65, for the evenly spaced LRV values we studied.
%% DHB: Note that this procedure is different than conditioning the statistical model on the target LRV.  We could switch to the latter if we wished.

@@

%% Figure condition 2: Excitations Fail
\begin{figure}
\centering
    \begin{subfigure}[b]{0.22\textwidth}   
        \caption{LRV estimates (excitations)}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure11/Figure11_a.pdf}
        \label{fig:case2IsomerizationEstimates}
    \end{subfigure}
        \begin{subfigure}[b]{0.22 \textwidth}
        \caption{RF responses (excitations)}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure11/Figure11_b.pdf}
        \label{fig:case2RFResponseIsomer}
    \end{subfigure}
            \begin{subfigure}[b]{0.22 \textwidth}
        \caption{LRV estimates (Normalized contrast)}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure11/Figure11_c.pdf}
        \label{fig:case2ContrastEstimates}
    \end{subfigure}
            \begin{subfigure}[b]{0.22 \textwidth}
        \caption{RF responses (Normalized contrast)}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure11/Figure11_d.pdf}
        \label{fig:case2RFResponseContrast}
    \end{subfigure}
    \caption{{\bf Condition 2 results:} (a) Mean estimated LRV $\pm$ 1 standard deviation for images in the test set obtained using linear regression and AMA estimators, based on the cone excitations. The relative RMSE is high (41.4\% for linear regression and 31.3\% for AMA based estimation). (b) Cone isomerizations projected on AMA receptive fields calculated using cone excitations. (c) Mean estimated LRV $\pm$ 1 standard deviation for images in the test set obtained using linear regression and AMA estimators, based on the contrast normalized cone excitations. The relative RMSE of estimation is 9.3\% for linear regression and 1.1\% for AMA based estimator. (d) The contrast normalized cone excitations projected along the first two AMA receptive fields. The receptive fields were learned using AMA and the contrast normalized cone excitations. For Condition 2, the response corresponding to individual image patches at each LRV level separates much better when the input is contrast normalized cone excitations rather than the cone excitations themselves.}
\label{fig:Condition2}
\end{figure}
%% DHB: The labels above each panel wrap in unfortunate ways. Fix.

@@

Condition 2 introduces variation in the spectral power distribution of the light sources.
This variation makes the task more difficult, because it causes more variation in cone excitations due to factors other than the target object LRV.
Figure~\ref{fig:Condition2}a and b show the results when we learned and evaluated the performance of AMA on the cone excitations.
Performance is poor.
Indeed, on average the estimates deviate considerably from the true LRV, as seen by the fact that the mean estimate
(red line) does not lie along the positive diagonal and by the fact that there is large estimation variability for each
value of the true LRV (as indicated by the width of the shaded red region).
For this case, linear regression is also very poor. Indeed, the linear regression estimates are essentially the same as would be obtained
by simply guessing the mean LRV of the training set (0.4).

Recall that there are two qualitatively distinct aspects to the variation in light source spectra.
One is a variation in relative spectral power distribution and the
other is a variation in overall intensity.
To separate the effect of these two aspects, we trained and evaluated AMA on the contrast normalized cone excitations.
Figure~\ref{fig:Condition2}(c, d) shows that performance improves greatly, and AMA performance is essentially perfect.
This observation is consistent with a large literature on computational color constancy that shows that when only illumination spectra are varied,
contrast-based representations provide an effective vehicle for achieving color constancy (REFS).
That same literature, however, shows that
contrast-based representations are less effective at supporting constancy when the spectra of background objects in the scene vary.
%% DHB: Need to add REFS in paragraph above.

%% Figure 12: Results for Condition 3
\begin{figure}
\centering
            \begin{subfigure}[b]{0.25 \textwidth}
        \caption{LRV estimates}
        \includegraphics[width=\textwidth, trim={0 0 0 1.3cm},clip]{../FiguresDraft4/Figure12/Figure12_a.pdf}
        \label{fig:case3Estimates}
    \end{subfigure} 
        \begin{subfigure}[b]{0.26\textwidth}
        \caption{Receptive field response}
        \includegraphics[width=\textwidth, trim={0 3mm 0 15mm},clip]{../FiguresDraft4/Figure12/Figure12_b.pdf}
        \label{fig:case3RFResponse}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.28 \textwidth}
	\caption{AMA receptive fields}
	\includegraphics[width=1.07\textwidth, trim={0 -0.6cm 0 -0.5cm}]{../FiguresDraft4/Figure12/Figure12_c.pdf}
	\label{fig:case3RFs}
    \end{subfigure}
    \caption{{\bf Condition 3 results:} (a) Mean estimated LRV $\pm$ 1 standard deviation for images in the test set obtained using linear regression and AMA estimators, based on the contrast normalized cone excitations. The relative RMSE is 22.9\%  for linear regression and 13.0\% for AMA. (b) Contrast normalized cone excitations for the training set projected along the first two AMA receptive fields. (c) First two AMA receptive fields learnt using the contrast normalized cone excitations.}
\label{fig:Condition3}
\end{figure}
%% Panel b shows training set projections, right?

@@

Our use of virtual images relies on datasets that characterize the statistical variation
in object reflectance and natural illumination spectra.
We believe these datasets provide reasonable approximations.
None-the-less, they can be extended and improved.
For example, the dataset of illuminations does not include the artificial illuminations
that we now live much of our lives under.
The surface reflectance datasets capture constraints on naturally occurring reflectance
functions, but the surfaces in the datasets were not sample the environment in a manner that
characterizes the relative frequency of different reflectance spectra in natural viewing.
Future work with virtual image datasets will benefit from targeted efforts to characterize
the natural statistics of the scene components that are varied.
Under the assumption that combination of different scene factors (e.g. surface reflectances
and illuminations) is independent, the virtual image approach modularizes the study
of variation in natural scenes.
%Supervised learning requires large labeled data sets.  Such data sets are not readily available for the study of color constancy. Although there are databases of calibrated color images, these do not provide ground truth information about surface reflectance and illuminantion at each image location \cite{ChakrabartiHyperspectral,NascimentoFoster2016,ParragaHyperspectralData,TkacikUpennHypersepctralData,skauli2013collection,olmos2004biologically}. There are some smaller databases consisting of images of posed scenes where object surface reflectances are measured individually (\citeNP{funt1988color,ciurea2003large}; \citeauthor{davidPennHyperspectral}), but these are not large enough to drive supervised learning.

@@

% MOVE THIS TO DISCUSSION
%This general approach has been adopted by other labs for the study of computational estimation of optical flow (Baker et al., 2011) and is becoming increasingly popular (Kovacs, Bell, Snavely, & Bala, 2017), as well as for the generation of stimuli for the study of human perception (e.g., Boyaci, Maloney, & Hersh, 2003; Todd & Mingolla, 1983; Arend & Reeves, 1986; Johnston, Cumming, & Parker, 1993). Of course, success within the domain of graphics images does not guarantee smooth generalization to real natural images, a point we return to in the Discussion.
% \cite{baker2011database}  \cite{kovacs17shading} \citeNP{boyaci2003effect, todd1983perception, arend1986Simultaneous, johnston1993integration}
% Autonmous vehicle work - ask BAW

>>>>

Early work on computational color constancy considered a simplified case with multiple flat matte objects and a single spatially diffuse illuminant \cite{LandRetinex,Buchsbaum80,MaloneyWandell86}.
Subsequent computational work considered more complex geometries and incorporated probabilistic descriptions of the statistics of naturally occurring scenes \cite{funt1988color, D'ZmuraConstancy3, barron2012color, D'ZmuraIversonSinger,BrainardFreeman}.
A key insight from this computational work is that stable color descriptors of a target object cannot be obtained from the light reflected from the target object alone.
But, it is possible to obtain such descriptors by analyzing jointly the light reflected from all of the objects in the scene.
As a consequence, the performance of color constancy algorithms will be affected by variation in the surface reflectance of all of the objects in the scene \cite{BrainardWandellRetinex}, not just by variation in the illuminant; evaluation of constancy algorithms must include assessment of their performance with respect to both variation in the illumination and the surface reflectance of all of the scene objects \cite{BrainardWandellRetinex,BrainardFreeman}.

@@

\subsection{Labeled training data} \label{method:VirtualWorld}
\subsubsection{Virtual naturalistic scenes}
The light that reflects from objects to the eyes depends on many factors.
These include the surface reflectance, texture, material and geometry of the objects, 
the spectral power distribution function of the illuminants, and the position of the observer.
We have developed a rendering package 
(\href{https://github.com/BrainardLab/VirtualWorldColorConstancy}{https://github.com/BrainardLab/VirtualWorldColorConstancy}) 
that allows us to construct models of naturalistic scenes, with key scene factors under programmatic control.
The package builds on our RenderToolblox4 package (\href{http://rendertoolbox.org}{http://rendertoolbox.org}) \cite{heasly2014rendertoolbox3},
and harnesses the open-source computer graphics renderer Mitsuba (\href{https://www.mitsuba-renderer.org}{https://www.mitsuba-renderer.org}, 
\citeNP{jakob2015mitsuba}) to produce physically-accurate images from the scene models.
Because each image is rendered from a known scene model, each image pixel can be labeled with 
the surface reflectance of the corresponding scene object.
By incorporating statistical models of the variation of object surface reflectance and daylight illumination, 
the pipeline allows us to produce large labeled sets of images that capture key aspects of the task-relevant 
statistical structure of natural scenes.

@@

For the primary results, we used the ``{\it Library}'' base scene and a spherical target object.
The library base scene contains 2 area lights. 
We inserted one additional spherical spherical light source into the scene.
The position and size of the inserted object, inserted light source and viewpoint were held fixed across all 
scenes in the database. The geometry is shown in Figure~\ref{fig:3DScene}.
Surface and illuminant spectra were drawn according to statistical models of naturally occurring spectra, as described below.
Multispectral images ($320 \times 240$ pixels) were rendered at 31 evenly-spaced wavelengths between $400$nm and $700$nm.
\subsubsection{Illumination spectra}
% Figure 6
\begin{figure}
\centering
    \begin{subfigure}[b]{0.24 \textwidth}
    \centering
	\caption{Granada dataset}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_a.pdf}
        \label{fig:granadaData}
    \end{subfigure}
	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{Statistical model}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_b.pdf}
        \label{fig:illuminantSamples}
    \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{CIE xy chromaticity}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_c.pdf}
        \label{fig:xyDiagram}
        \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{Color swatches}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_d.pdf}
        \label{fig:sRGBIlluminant}
    \end{subfigure}
    \caption{{\bf Statistical model of illumination spectra:} (a) Normalized Granada dataset. Each spectrum is normalized by its mean (with mean taken over wavelength). (b) Sample spectra generated using the statistical model derived from normalized Granada dataset. (c) CIE xy chromaticities of the Granda dataset (black) and the samples shown in b (red). F(d) sRGB renditions of the samples shown in b.}
\label{fig:illuminant}
\end{figure}
%% DHB: The chromaticities in c and renderings in d are for the spectra in b, right?  I wrote the caption to say so, just checking.
%% VS: Yes

@@

Because our multivariate Gaussian model is based on normalized spectra, it does not embody the large variation in overall intensity of natural daylights.
To obtain illuminant spectra with intensity variation similar to the Granada dataset, we scaled each randomly generated relative spectrum.
The scale factors were sampled randomly on a log-uniform scale in the range [0.15, 150].
%% DHB: Check that these values are right.  Linear spacing or log spacing?
%% VS: Yes, log spacing, I have updated the text.

\subsubsection{Surface reflectance spectra}
% Figure 7 Surface Reflectance Data and Model
\begin{figure}
\centering
	\begin{subfigure}{0.24 \textwidth}
    \centering
            \caption{Munsell and Vrhel dataset}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_a.pdf}
        \label{fig:reflectanceSpectra}
    \end{subfigure}
    \begin{subfigure}{0.24 \textwidth}
    \centering
        \caption{Statistical model}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_b.pdf}
        \label{fig:reflectanceSamples}
    \end{subfigure}
    \begin{subfigure}{0.24 \textwidth}
    \centering
    \caption{CIE xy chromaticity}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_c.pdf}
        \label{fig:xyChroReflectance}
    \end{subfigure}    
    \centering
	\begin{subfigure}{0.24 \textwidth}
    \centering
        \caption{Color swatches}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_d.pdf}
        \label{fig:backgroundSwatches}
    \end{subfigure}
    \caption{{\bf Statistical model of surface reflectance:} (a) Surface reflectance spectra from the Munsell and Vrhel datasets. (b) Sample spectra generated using the surface reflectance statistical model. (c) CIE xy chromaticity diagram of the Munsell and Vrhel reflectances and samples shown in b (red). (d) sRGB renditions of the samples shown in b, rendered under the CIE D65 daylight.}
\label{fig:surfaceReflectanceGeneration}
\end{figure}
%% DHB: The chromaticities in c and renderings in d are for the spectra in b, right?  I wrote the caption to say so, just checking.
%% VS: Yes

@@

For generating the target object reflectance at a particular luminance, the values in a generated spectrum were 
scaled such that the LRV had the desired value (see appendix).
Figure \ref{fig:targetSwatches} shows color renderings of target reflectance spectra under CIE illuminant D65, for the evenly spaced LRV values we studied.
%% DHB: Note that this procedure is different than conditioning the statistical model on the target LRV.  We could switch to the latter if we wished.
%% VS: I don't think it matters if we care about keeping only one color co-ordinate fixed.

@@

%% Figure condition 2: Excitations Fail
\begin{figure}
\centering
    \begin{subfigure}[b]{0.22\textwidth}   
        \caption{\centering{LRV estimates} \newline\centering{(Excitations)}}
        \includegraphics[width=\textwidth, trim={0 0 0 1.5cm},clip]{../FiguresDraft4/Figure11/Figure11_a.pdf}
        \label{fig:case2IsomerizationEstimates}
    \end{subfigure}
        \begin{subfigure}[b]{0.22 \textwidth}
        \caption{\centering{RF responses} \newline\centering{(Excitations)}}
        \includegraphics[width=\textwidth, trim={0 0 0 1.5cm},clip]{../FiguresDraft4/Figure11/Figure11_b.pdf}
        \label{fig:case2RFResponseIsomer}
    \end{subfigure}
            \begin{subfigure}[b]{0.22 \textwidth}
        \caption{\centering{LRV estimates}  \newline \centering{(Normalized contrast)}}
        \includegraphics[width=\textwidth, trim={0 0 0 1.5cm},clip]{../FiguresDraft4/Figure11/Figure11_c.pdf}
        \label{fig:case2ContrastEstimates}
    \end{subfigure}
            \begin{subfigure}[b]{0.22 \textwidth}
        \caption{\centering{RF responses}  \newline \centering{(Normalized contrast)}}
        \includegraphics[width=\textwidth, trim={0 0 0 1.5cm},clip]{../FiguresDraft4/Figure11/Figure11_d.pdf}
        \label{fig:case2RFResponseContrast}
    \end{subfigure}
    \caption{{\bf Condition 2 results:} (a) Mean estimated LRV $\pm$ 1 standard deviation for images in the test set obtained using linear regression and AMA estimators, based on the cone excitations. The relative RMSE is high (41.4\% for linear regression and 31.3\% for AMA based estimation). (b) Cone isomerizations projected on AMA receptive fields calculated using cone excitations. (c) Mean estimated LRV $\pm$ 1 standard deviation for images in the test set obtained using linear regression and AMA estimators, based on the contrast normalized cone excitations. The relative RMSE of estimation is 9.3\% for linear regression and 1.1\% for AMA based estimator. (d) The contrast normalized cone excitations projected along the first two AMA receptive fields. The receptive fields were learned using AMA and the contrast normalized cone excitations. For Condition 2, the response corresponding to individual image patches at each LRV level separates much better when the input is contrast normalized cone excitations rather than the cone excitations themselves.}
\label{fig:Condition2}
\end{figure}

@@

Condition 2 introduces variation in the spectral power distribution of the light sources.
This variation makes the task more difficult, because it causes more variation in cone excitations due to factors other than the target object LRV.
Figure~\ref{fig:Condition2}(a,b) show the results when we learned and evaluated the performance of AMA on the cone excitations.
Performance is poor.
Indeed, on average the estimates deviate considerably from the true LRV, as seen by the fact that the mean estimate
(red line) does not lie along the positive diagonal and by the fact that there is large estimation variability for each
value of the true LRV (as indicated by the width of the shaded red region).
For this case, linear regression is also very poor. Indeed, the linear regression estimates are essentially the same as would be obtained
by simply guessing the mean LRV of the training set (0.4).

Recall that there are two qualitatively distinct aspects to the variation in light source spectra.
One is a variation in relative spectral power distribution and the
other is a variation in overall intensity.
To separate the effect of these two aspects, we trained and evaluated AMA on the contrast normalized cone excitations.
Figure~\ref{fig:Condition2}(c,d) shows that performance improves greatly, and AMA performance is essentially perfect.
This observation is consistent with a large literature on computational color constancy that shows that when only illumination spectra are varied,
contrast-based representations provide an effective vehicle for achieving color constancy (REFS).
That same literature, however, shows that
contrast-based representations are less effective at supporting constancy when the spectra of background objects in the scene vary.
%% DHB: Need to add REFS in paragraph above.

%% Figure 12: Results for Condition 3
\begin{figure}
\centering
            \begin{subfigure}[b]{0.25 \textwidth}
        \caption{LRV estimates}
        \includegraphics[width=\textwidth, trim={0 0 0 1.3cm},clip]{../FiguresDraft4/Figure12/Figure12_a.pdf}
        \label{fig:case3Estimates}
    \end{subfigure} 
        \begin{subfigure}[b]{0.26\textwidth}
        \caption{Receptive field response}
        \includegraphics[width=\textwidth, trim={0 3mm 0 15mm},clip]{../FiguresDraft4/Figure12/Figure12_b.pdf}
        \label{fig:case3RFResponse}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.28 \textwidth}
	\caption{AMA receptive fields}
	\includegraphics[width=1.07\textwidth, trim={0 -0.6cm 0 -0.5cm}]{../FiguresDraft4/Figure12/Figure12_c.pdf}
	\label{fig:case3RFs}
    \end{subfigure}
    \caption{{\bf Condition 3 results:} (a) Mean estimated LRV $\pm$ 1 standard deviation for images in the test set obtained using linear regression and AMA estimators, based on the contrast normalized cone excitations. The relative RMSE is 22.9\%  for linear regression and 13.0\% for AMA. (b) Contrast normalized cone excitations for the training set projected along the first two AMA receptive fields. (c) First two AMA receptive fields learnt using the contrast normalized cone excitations.}
\label{fig:Condition3}
\end{figure}
%% Panel b shows training set projections, right?
%% VS: Yes

@@

Our use of virtual images relies on datasets that characterize the statistical variation
in object reflectance and natural illumination spectra.
We believe these datasets provide reasonable approximations.
None-the-less, they can be extended and improved.
For example, the dataset of illuminations does not include the artificial illuminations
that we now live much of our lives under.
The surface reflectance datasets capture constraints on naturally occurring reflectance
functions, but the surfaces in the datasets were not sample the environment in a manner that
characterizes the relative frequency of different reflectance spectra in natural viewing.
Future work with virtual image datasets will benefit from targeted efforts to characterize
the natural statistics of the scene components that are varied.
Under the assumption that combination of different scene factors (e.g. surface reflectances
and illuminations) is independent, the virtual image approach modularizes the study
of variation in natural scenes.

@@
